### 2025-10-23 bv-triples-main.tex

%+----------------------------------------------------------+
%| SLIDES: Construction and Reduction of the L-infinity Algebra of Observables
%|         Associated with a BV-Module
%| Contents: - 60 minutes (estimated duration ~2.5 minutes per slide, ~24 slides)
%|
%| Author: Antonio Miti
%| Event: ARTS — Algebra and Representation Theory Seminar
%| Place: Università di Roma  "Tor Vergata "
%| Date: 21/10/2025
%| Link: https://www.mat.uniroma2.it/~ricerca/algebr/page-AREA_files/page-AREA-2_files/talks.html
%+----------------------------------------------------------+

%- HandOut Flag -----------------------------------------------------------------------------------------
	\newif\ifHandout
	\Handouttrue  %uncomment for the printable version
	%Handling of flags it is not preserved when passing to standalone-subfiles!

%- D0cum3nt ----------------------------------------------------------------------------------------------
\ifHandout
	\documentclass[handout,10pt]{beamer}   
	\setbeameroption{show notes} %print notes   
\else
	\documentclass[10pt]{beamer}
\fi

%- Packages ----------------------------------------------------------------------------------------------
\usepackage{custom-style}
\usepackage{math}

%- Bibliography (Biber) ----------------------------------------------------------------------------------
\usepackage[backend=biber,style=alphabetic,maxnames=2]{biblatex}
\bibliography{bibfile.bib}

%- L30's ----------------------------------------------------------------------------------------------
\usepackage{quiver} %https://q.uiver.app/ % Not in Miktex!
\usetikzlibrary{nfold}
\usetikzlibrary{decorations.pathmorphing} 
\renewcommand{\Ham}{\mathrm{Ham}}
\usepackage{enumitem}
\setlist{topsep=0pt, leftmargin=*,label=$\bullet$}

%--Beamer Style-----------------------------------------------------------------------------------------------
\usetheme{toninus}

\usetikzlibrary{backgrounds}
  \tikzset{
    invisible/.style={opacity=0},
    visible on/.style={alt=#1{}{invisible}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
    },
  }

%- T1tle P4g3 -------------------------------------------------------------------------------------------
\title{Construction and Reduction of the $L_\infty$ Algebra of Observables Associated with a BV-Module} 
\subtitle{}
\author[AMM]{
	\href{https://www.antoniomiti.it/}{Antonio Michele Miti} \\
	{(\small joint work w/ \href{https://www.ryvkin.eu}{Leonid Ryvkin})}
}
\institute[SUR]{
	Sapienza Università di Roma \\
	Rome, Italy 	\\
	\vspace{.5em}
  \begin{tabular}[h]{ccc}
      \href{https://www.mat.uniroma1.it/en}{\includegraphics[width=6cm]{./Logos/Sur_logo}} & & 
      \href{https://civis3i.univ-amu.fr/en/civis3i-alliance-programme}{\includegraphics[width=3cm]{Logos/Civis_logo}}
  \end{tabular}    
}
\date[TorV25] % (optional, should be abbreviation of conference name)
{

	{\vskip 1ex}
	\href{https://www.mat.uniroma2.it/~ricerca/algebr/page-AREA_files/page-AREA-2_files/talks.html}{- ARTS - Algebra and Representation Theory Seminar in Roma "Tor Vergata"}
	\\
	Universit\'a di Roma - Tor Vergata, October 2025
}

%+----------------------------------------------------------+
%- D0cum3nt ------------------------------------------------+
%+----------------------------------------------------------+
\begin{document}
%+----------------------------------------------------------+

%----------------------------------------------------------------------------------+
\begin{frame}  % Alternative: \maketitle outside of frame
	\titlepage
	\ifHandout
		\tikz[overlay,remember picture]
		{
	    	%	\node at ($(current page.west)+(1.5,0)$) [rotate=90] {\Huge\textcolor{gray}{\today}};
	    	\node[        
	    		draw,
	    		shape border rotate=90,
			isosceles triangle,
			isosceles triangle apex angle=90,
			fill=yellow]
	        		at ($(current page.north east)-(1,1)$) [rotate=-45] {\textcolor{red}{Handout version}};
		}
	\fi
\end{frame}
\addtocounter{framenumber}{-1}
\note{
	%\textbf{\underline{OUTLINE}}:
	%\tableofcontents
	\textbf{Abstract:}
	\\
	\footnotesize
Multisymplectic manifolds generalize symplectic manifolds by carrying a closed, nondegenerate differential form of degree higher than 2. Such structures naturally arise in the geometric formulation of classical field theories.
Rogers (2010) showed that, just as a symplectic manifold yields a Poisson algebra of functions, an n-plectic manifold gives rise to an n-term Lie infinity algebra of observables.
Remarkably, his construction is purely algebraic and based only on Cartan calculus, suggesting that this higher “Poisson algebra” can be generalized beyond manifolds.
In this talk, we present such a generalization within the setting of Gerstenhaber algebras and Batalin–Vilkovisky (BV) modules, which provide an algebraic formulation of Cartan calculus relevant to noncommutative geometry.
This framework allows us to define Lie infinity algebras of observables in a fully algebraic way, without referring to an underlying manifold.
As an application, we address the reduction of multisymplectic observables in the presence of constraints or symmetries. Building on the work of Dippel, Esposito, and Waldmann on “constraint triples” for coisotropic reduction, we adapt this formalism to BV-modules and their Lie infinity algebras.
This yields a conceptual setting for the algebraic reduction of multisymplectic observables, as developed in our joint work with Casey Blacker (SIGMA 2024).
The results are part of a collaboration with Leonid Ryvkin, published in *Differential Geometry and its Applications* (2025).

}
%--------------------------------------------------------------------------------------------------

%+----------------------------------------------------------+
\section{Introduction}
%+----------------------------------------------------------+
	\input{bv-triples-intro}
%+----------------------------------------------------------+

%+----------------------------------------------------------+
\section{BV-Modules: Algebraic Cartan Calculus}
%+----------------------------------------------------------+
	\input{bv-triples-background}
%-----------------------------------------------------------%

%+----------------------------------------------------------+
\section{$L_\infty$ Algebra Associated with a BV-Module}
%+----------------------------------------------------------+
	\input{bv-triples-Rogers-generalized}
%-----------------------------------------------------------%

%+----------------------------------------------------------+
\section{An Application: Multisymplectic Reduction}
%+----------------------------------------------------------+
	\input{bv-triples-reduction}
%-----------------------------------------------------------%

%--------------------------------------------------------------------------------------------------%
	\addtocounter{framenumber}{-1}
		\begin{frame}{Conclusions}
			\begin{upshotblocktitle}[Algebraic formalism for $L_\infty$ observables]
				We generalized Rogers' construction to constraint BV-modules.
			\end{upshotblocktitle}
			\vfill\pause

			\begin{upshotblocktitle}[Recovered previous reduction procedures]
				 Emploied this machinery to provide a conceptual framework for algebraic multisymplectic reduction of observables via constraint triples 
				 \smallskip

				 \fullcite{Blacker-M-Ryvkin-2024}\nocite{Blacker-M-Ryvkin-2024}
			\end{upshotblocktitle}
			\vfill\pause

			\begin{outlookblock}[An operadic reformulation]
				Rather than defining suitable constraint triples "by hand", it could be interesting to provide a systematic treatment, e.g. within the language of operads.
			\end{outlookblock}
			\vfill\pause

		  	\centering 
		  	{\Huge\color{red} 
		  	\emph{Thank you for your attention!}\normalsize \color{black}}
			\vfill
			%

		\end{frame}
		\note[itemize]{
			\item 				The constraint triples formalism provides a powerful tool to treat singular geometric (and algebraic) settings. 
		}
%------------------------------------------------------------------------------------------------

%-------------------------------------------------------+
% APPENDIX
%-------------------------------------------------------+
%-------------------------------------------------------+
\appendix
%-------------------------------------------------------+
%-------------------------------------------------------+
\section{Complementary Material}
	\input{bv-triples-complementary-material}
%------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
\section{References \& Aknowledgements}
%------------------------------------------------------------------------------------------------
	\input{bv-triples-acknowledgements}

%+----------------------------------------------------------+
\end{document}
%+----------------------------------------------------------+